\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb}
\usepackage{microtype}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{tikz}
\usetikzlibrary{positioning, fit, arrows.meta, calc, backgrounds, shapes.geometric}
\usepackage[round]{natbib}
\usepackage{hyperref}
\hypersetup{colorlinks=true, linkcolor=black, urlcolor=blue}

% Number this section as §3 to match the larger paper.
\setcounter{section}{2}

% Convenience macros.
\newcommand{\anchorW}{W_{\text{anchor}}}
\newcommand{\anchorB}{b_{\text{anchor}}}
\newcommand{\Lewc}{L_{\text{EWC}}}
\newcommand{\Ltask}{L_{\text{task}}}
\newcommand{\reff}{r_{\text{eff}}}
\newcommand{\ac}{\text{ac}}
\newcommand{\R}{\mathbb{R}}

\title{Trioron --- Methods (\S3)}
\author{}
\date{}

\begin{document}
\maketitle

\section{Methods}
\label{sec:methods}

\paragraph{Notation.}
$n$ is the number of nodes in a layer; $d$ is the layer's incoming
dimension; $B$ is batch size; $t$ indexes tasks in a curriculum.

\subsection{The Trioron Node}
\label{sec:trioron-node}

A \emph{trioron} is a single computational unit characterized by three
coupled state variables, generalizing the standard artificial neuron:

\begin{itemize}[leftmargin=1.5em]
  \item $w \in \R^{d}$ --- incoming weights (a row of the layer's weight
        matrix).
  \item $\lambda \in \R_{\geq 0}$ --- per-node \emph{plasticity coefficient}:
        a per-unit scalar that scales the unit's contribution to the
        elastic-weight consolidation (EWC) penalty. Larger
        $\lambda \Rightarrow$ stiffer; smaller $\lambda \Rightarrow$ more
        plastic.
  \item $u \in \R$ --- per-node \emph{utility score}: a running estimate
        of the unit's contribution to the network's outputs. Used as the
        signal for pruning decisions (cells with persistently low $u$ are
        candidates for removal).
\end{itemize}

A \emph{trioron layer} aggregates $n$ trioron nodes that share an incoming
dimension. We hold the per-node state in vectors $\lambda, u \in \R^{n}$
and a weight matrix $W \in \R^{n \times d}$ where row $i$ is $w_i$. The
forward pass for input batch $X \in \R^{B \times d}$ is
\begin{equation}
  H = \sigma\!\bigl(X\,(r \odot W)^{\top} + b\bigr),
\end{equation}
where $\sigma$ is the activation function, $b \in \R^{n}$ is the bias, and
$r \in [0, 1]^{n}$ is the layer's \emph{routing scale} --- a per-node
multiplicative gain on incoming weights, defaulting to all ones. The
routing scale is the substrate for routing-starvation consolidation
(\S\ref{sec:dreaming-phase}).

Each layer additionally maintains EWC state: anchor weights $\anchorW$ and
$\anchorB$ snapshotted at task boundaries, and Fisher-information
accumulators $F_W, F_b$ updated as a running EMA of squared gradients.
Per-node $\lambda$ is derived from the row-mean of Fisher:
$\lambda_i = \operatorname{mean}_j F_{W,i,j}$.

\paragraph{Epigenetic analogue.}
$w$ is the synaptic-strength channel (modulated by experience-dependent
LTP/LTD); $\lambda$ is a per-cell modifier on plasticity, with biological
correlates including DNA-methylation status of plasticity-related genes
(e.g., BDNF) and perineuronal-net maturation, both of which gate how
readily a cell's synapses can be modified; $u$ is the cell-importance
signal that biological systems approximate via activity-dependent
neurotrophin release. The triplet $(w, \lambda, u)$ captures \emph{which
connections are present, how rigid they are, and how much the cell
matters} --- three quantities that biological systems modulate
independently.

\subsection{Continual-Learning Substrate (EWC with Per-Node Plasticity)}
\label{sec:ewc-substrate}

Following \citet{kirkpatrick2017overcoming}, we apply elastic weight consolidation
to protect prior-task knowledge. The penalty for a single trioron layer is
\begin{equation}
  \Lewc = \sum_i \lambda_i \left[
    \sum_j (W_{ij} - W_{\text{anchor},ij})^{2}
         + (b_i - b_{\text{anchor},i})^{2}
  \right].
\end{equation}
The total loss is $L = \Ltask + \beta \cdot \Lewc$ for an EWC strength
$\beta$ set per-curriculum. We refresh $\lambda_i$ from Fisher at the end
of each task (after a batched re-estimation pass over recent task data),
then snapshot $\anchorW \leftarrow W$ as the new anchor. This separation
--- estimate Fisher at task end, then anchor --- follows the original EWC
recipe and avoids the EMA-drift artifacts of online updating during
training.

For longer curricula we optionally accumulate Fisher across tasks via the
Online EWC variant \citep{schwarz2018progress}:
\begin{equation}
  F_t \leftarrow \gamma \cdot F_{t-1} + F_{\text{new}},
  \qquad \gamma \in [0, 1],
\end{equation}
with $\gamma = 1$ reducing to vanilla EWC. We use Online EWC
$\gamma = 0.95$ as a competitor baseline in \S4.

\subsection{Structural Plasticity}
\label{sec:structural-plasticity}

The architecture supports in-place growth and pruning of nodes during
training, with cross-layer consistency.

\paragraph{Cellular division (\texttt{grow\_node}).}
Adds one row to a layer's weight matrix, extends the per-node state
vectors $\lambda, u, r$ by one entry, and (if a downstream layer exists)
extends that layer's incoming dimension by one column. The new node is
initialized fully plastic ($\lambda_{\text{new}} = 0$,
$u_{\text{new}} = 0$, $r_{\text{new}} = 1$) so it can adapt freely while
existing nodes remain protected by EWC. The new row's incoming weights are
initialized along the top principal direction of the residual signal
$D = f(X_a) - f(X_b)$ at the layer below --- the direction of unmet
representational variance, computed by SVD over a probe batch.

\paragraph{Cellular pruning (\texttt{prune\_node}).}
Removes a node's row, contracts the state vectors, and drops the
corresponding input column on the next layer. By default the pruned node's
outgoing column is \emph{redistributed} to its weight-cosine-nearest peer
on the next layer before removal, preserving approximate input--output
behavior of the layer.

\paragraph{Growth trigger.}
A node is grown only when three conditions are jointly satisfied over a
window of $W$ steps:
\begin{enumerate}[leftmargin=2em]
  \item \emph{Loss plateau.} The loss has not improved by more than
        $\varepsilon_{\text{loss}}$ relative to the prior window.
  \item \emph{Rank saturation.} The effective rank
        \begin{equation*}
          \reff = \exp\!\left(-\sum_k p_k \log p_k\right),
          \qquad p_k = \sigma_k \big/ \textstyle\sum_k \sigma_k,
        \end{equation*}
        computed from the normalized singular values of the latent
        activation matrix, has approached the latent dimension
        ($\text{latent\_dim} - \reff < \varepsilon_{\text{rank}}$).
  \item \emph{Gradient stability.} The median gradient norm is bounded
        $g_{\min} \leq |\tilde g| \leq g_{\max}$ --- the optimizer is not
        exploding or vanishing.
\end{enumerate}
The conjunction of plateau-and-saturation-and-stability ensures that the
network grows only when it has \emph{actually run out of capacity for the
current task}, not merely when training is slow.

\paragraph{Resource ceilings.}
A pre-flight check rejects a proposed division if it would exceed a
per-curriculum byte budget \textit{cap\_bytes}. The cap can be lifted
between phases (we do this in the extension experiment of
\S\ref{sec:ship-wake-extend} and \S4.5). This implements \emph{resource
awareness} --- the network knows when not to grow even if growth would
help. Biologically analogous to metabolic limits on adult neurogenesis.

\paragraph{Epigenetic analogue.}
Growth corresponds to adult neurogenesis (which remains active in the
dentate gyrus and a few other regions throughout life); the
rank-saturation trigger reflects the biological observation that new
neurons are recruited preferentially when existing circuits saturate.
Pruning corresponds to developmental and ongoing synaptic pruning, with
cosine-nearest-peer redistribution as a soft analog of compensatory
hypertrophy.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
  >={Latex[length=2.2mm]},
  font=\small,
  layer/.style={draw, rounded corners=3pt, minimum width=22mm,
                minimum height=26mm, align=center, inner sep=3pt, fill=white},
  frozen/.style={layer, fill=gray!15},
  plastic/.style={layer, fill=blue!10},
  ann/.style={font=\scriptsize, align=center},
  arr/.style={->, thick, shorten >=1pt, shorten <=1pt}
]
\node[layer] (in) {\textbf{Input}\\[1pt]$\R^{784}$\\[1pt]
                   \scriptsize MNIST / FMNIST /\\\scriptsize EMNIST pixel};
\node[frozen, right=10mm of in] (l0) {\textbf{L0}\\[1pt]$\R^{128}$\\[1pt]
                   \scriptsize \emph{frozen Gaussian}\\
                   \scriptsize random projection};
\node[plastic, right=10mm of l0] (l1) {\textbf{L1}\\[1pt]
                   \scriptsize plastic\\$32 \to$ cap\\
                   \scriptsize ($\le 8\text{K}$ params)};
\node[plastic, right=10mm of l1] (hd) {\textbf{Head}\\[1pt]
                   \scriptsize plastic\\$|C_t|$ classes\\
                   \scriptsize +1 col / new class};
\draw[arr] (in) -- (l0);
\draw[arr] (l0) -- (l1);
\draw[arr] (l1) -- (hd);
\node[ann, above=7mm of l1] (gtrig) {plateau $\wedge$ rank-sat $\wedge$ stable-grad\\
                                     $\Rightarrow$ \texttt{grow\_node}};
\draw[->, dashed, thick, green!50!black] (gtrig.south) -- (l1.north);
\node[ann, below=7mm of l1] (purge) {dream-time \texttt{purge\_node}\\
                                     (sRNA-capped, $N_{\max}$ per cycle)};
\draw[->, dotted, thick, red!60!black] (purge.north) -- (l1.south);
\node[ann, below=7mm of hd] (ewc) {EWC: per-node $\lambda_i$,\\
                                   anchor $\anchorW$};
\draw[densely dotted] (ewc.north) -- (hd.south);
\draw[densely dotted] (ewc.north west) to[bend right=8] (l1.south east);
\end{tikzpicture}
\caption{Base trioron architecture for chained-15. The frozen Gaussian
L0 projection provides a stable code space for manifold replay
(\S\ref{sec:manifold-replay}). The plastic L1 grows under the
conjunction of loss-plateau, rank-saturation, and gradient-stability
triggers up to a per-curriculum byte cap; the head grows one column per
new class. Dream-time purge removes low-utility nodes, gated by the
per-cycle sRNA cap (\S\ref{sec:srna-cap}). EWC anchors plus per-node
plasticity $\lambda_i$ pin prior-task knowledge on plastic layers.}
\label{fig:trioron-arch}
\end{figure}

\subsection{The Dreaming Phase}
\label{sec:dreaming-phase}

Between training tasks we run an offline \emph{dreaming block}: a short
sleep-time pass that consolidates what was just learned without exposing
the network to new data. EWC remains active inside the dream block so
that consolidation pulls weights toward their anchored state. The dream
block runs three sequential stages:
\begin{equation*}
  \text{replay} \;\to\; \text{purge} \;\to\; \text{archive}.
\end{equation*}
Replay (\S\ref{sec:replay}) is the load-bearing accuracy mechanism; the
dream archive (\S\ref{sec:archive}) is the load-bearing storage
mechanism; purge is gated by a per-cycle event cap
(\S\ref{sec:srna-cap}).

\subsubsection{Replay}
\label{sec:replay}

Replay rehearses past tasks under EWC. We support two replay sources:
\begin{itemize}[leftmargin=1.5em]
  \item \emph{Episodic exemplars (hippocampal buffer).} When real-data
        storage is permitted we maintain a buffer of $K$ samples per past
        task, sampled uniformly at task end. During replay the buffer is
        shuffled with the current task's batch (1{:}1 by default) and the
        cross-entropy loss is evaluated on both. $K \in \{1, 10, 20, 50\}$
        is the storage axis we sweep in \S4.3. $K = 0$ routes replay
        through the storage-free pseudo-rehearsal pathway of
        \S\ref{sec:manifold-replay}.
  \item \emph{Manifold pseudo-rehearsal (storage-free).} Per-class
        statistics of L0 activations are stored at task end and used to
        synthesize replay samples; see \S\ref{sec:manifold-replay} for the
        full mechanism. This is the headline rehearsal channel.
\end{itemize}
Conceptually analogous to sharp-wave ripples during slow-wave sleep, which
replay recent and remote memories at compressed timescales.

\subsubsection{The sRNA Cap}
\label{sec:srna-cap}

Each dream block bounds the number of structural-pruning events per
layer by a per-cycle cap $N_{\max}$. The cap holds even when many
candidates pass the utility threshold of \S\ref{sec:purge}: empirically
we observe across the dream-time mechanism design space we explored that
$\sigma$-distance from baselines degrades monotonically with structural
event count, motivating small caps. The default in the working
configuration is $N_{\max} = 1$ --- at most one purge event per layer
per dream cycle, allowing replay to absorb the structural change before
the next event drifts on top.

\paragraph{Epigenetic analogue.}
The cap is a direct analog of resource-limited small-RNA pools that gate
sleep-cycle synaptic homeostasis. miRNAs involved in synaptic remodeling
are produced and consumed in stoichiometric amounts during a single sleep
cycle; the cellular machinery cannot perform arbitrarily many remodeling
events per cycle even when many candidates are present. The ``fewer
events $=$ better'' empirical pattern reproduces the biological
observation that few consolidation events per cycle are protective.

\subsubsection{Purge}
\label{sec:purge}

The purge stage drops nodes whose utility $u_i$ falls below a threshold
$\tau_u$, subject to the per-layer cap $N_{\max}$ of
\S\ref{sec:srna-cap}. It reuses the structural pruning machinery of
\S\ref{sec:structural-plasticity}, including cross-layer fan-in cleanup
and cosine-nearest-peer redistribution. Distinct from the in-training
pruner (\S\ref{sec:structural-plasticity}): purge is a sleep-time
structural sweep with a strict event cap, not an event-clock check
during waking training.

\subsubsection{Pseudo-code of a Dream Block}
\label{sec:dream-pseudocode}

\begin{verbatim}
for each task t in curriculum:
    train_one_task(t)                         # waking
    consolidate_task(t)                       # Fisher update + anchor
    if dreaming_enabled:
        replay(buffer = hippo or manifold,
               steps = K_r,
               ewc_active = True)
        purge(threshold = tau_u,
              max_events_per_layer = N_max)   # sRNA cap
        archive_block(...)                    # see archive section
\end{verbatim}

\subsection{Manifold Replay (Storage-Free Pseudo-Rehearsal)}
\label{sec:manifold-replay}

The headline rehearsal mechanism is \emph{manifold replay}: a
trioron-native pseudo-rehearsal scheme that summarizes each past class as
a small set of statistics in the network's first-hidden-layer (L0)
representation rather than storing exemplars.

\paragraph{Storage.}
When a task $t$ ends we collect L0 activations for each class $c$ that
appeared in the task and store the per-class mean
$\mu_c \in \R^{L_0}$ and per-dimension standard deviation
$\sigma_c \in \R^{L_0}$ (a diagonal-Gaussian summary). Storage per class is
$2 \cdot L_0 \cdot 4 = 1024$ bytes at $L_0 = 128$. A 30-class run costs
$\approx\!30$~KB total --- independent of the dataset size and of the
number of samples per task.

\paragraph{Replay.}
During the dream block we draw pseudo-samples for each past class $c$:
\begin{equation}
  \tilde z = \mu_c + \sigma_c \odot \varepsilon,
  \qquad \varepsilon \sim \mathcal{N}(0, I),
\end{equation}
then forward them through the \emph{upper} layers via
\texttt{forward\_from\_layer(z\_tilde, start = 1)} to obtain logits.
Cross-entropy loss on $(\tilde z, c)$ is added to the total loss. Replay
is performed at the same EWC strength as waking training.

\paragraph{Why per-class statistics work where logit-inversion fails.}
Three single-seed inversion variants --- DeepInversion-style code
synthesis (one-shot, refresh-all, and diversity-regularized) --- all
peaked at $0.27$ full-softmax accuracy on chained-15. Logit inversion
locates class-$c$ attractor \emph{peaks}, not class-$c$ manifold
\emph{samples}; the resulting points concentrate at adversarial extrema
that do not reflect actual class structure. This is the same pathology
that defeated earlier engram-replay variants (gradient-ascent on real-mean
seeds; L1-feature distillation; chain combinations of engram with
differential L1 storage). The $(\mu, \sigma)$ summary lives on the real L0
manifold by construction; inversion attractors do not. An ablation
removing per-class $\sigma$ ($\mu$-only replay) loses $\approx\!0.07$
absolute on full-softmax at 30 classes (single-seed), confirming that
per-class noise scale is load-bearing --- roughly $12\%$ of manifold
replay's advantage.

\paragraph{Why L0 is frozen.}
Manifold replay assumes that L0's encoding is approximately stationary
across the curriculum, so that statistics collected after task $t$ remain
valid at task $t + k$. We enforce this by freezing L0 at initialization to
a fixed random Gaussian projection (weights drawn from
$\mathcal{N}(0, 2/d)$ at init and never updated). A separation-probe
diagnostic on the chained-15 curriculum gives a class-prototype
discriminability ratio of $1.063$ in the frozen L0 space --- modest but
non-trivial, with Fashion-MNIST contributing the lowest separability
($1.04$) as expected from its irreducible within-domain confusions. The
ratio rules out a single-prototype scheme (L0-space class means alone are
not separable enough) and motivates the stored-$(\mu, \sigma)$ sketch as a
manifold summary rather than a class-prototype classifier.

\paragraph{Epigenetic analogue.}
Per-class $(\mu, \sigma)$ is a compact ``engram fingerprint'' --- an
offline sketch of the data manifold rather than a stored sample.
Functionally closer to consolidated memory traces (gist-level recall) than
to episodic exemplar storage (which the hippocampal-buffer baseline of
\S\ref{sec:replay} implements explicitly).

\subsection{The Dream Archive}
\label{sec:archive}

Long curricula accumulate parameters that should not keep moving --- rows
whose Fisher has plateaued and whose contribution is stable across recent
tasks. We expose these as a \emph{dream archive}: a two-phase mechanism
for locking and quantizing such rows.

\paragraph{Phase 1 --- mid-curriculum row archive.}
After each \texttt{consolidate\_task}, every row of every layer is scored
along three axes: (a) $\lambda$ percentile (top
\textit{archive\_lam\_top\_percentile} of the layer), (b)
gradient-magnitude ceiling (median $|g|$ over the recent window below
\textit{archive\_grad\_mag\_floor}), and (c) apoptosis-pulse bound
($p \leq$ \textit{archive\_pulse\_max} --- never archive a row recently
pulsed by an apoptosis event). A row that satisfies all three for
\textit{archive\_streak\_threshold} consecutive tasks is \emph{archived}:
a per-row gradient mask is set so subsequent backward passes leave the row
at its anchor value. The mask is enforced via
\texttt{mask\_archived\_grads\_all()} between \texttt{loss.backward()} and
\texttt{opt.step()} (sequenced after PackNet/HAT grad surgery on
competitor arms). A per-layer cap \textit{archive\_max\_per\_layer}
prevents any one layer from being fully archived in a single pass.

\paragraph{Phase 2 --- end-of-curriculum int8 quantization.}
At end of training we snap each archived row to int8 with a per-row
symmetric scale ($8$~bits/weight plus one fp32 scale per row). Active
(non-archived) weights are rounded to BF16 precision. The Phase 2 path is
implemented as a snapshot--quantize--re-evaluate--restore sequence in the
bench so we can report storage and accuracy numbers from the same trained
checkpoint; the deployment path applies the snap permanently.

\paragraph{Storage accounting.}
With BF16 as the deployment baseline, archive $+$ int8 reduces network
weight storage by $\approx\!40\%$ across all arms ($213.7 \to 127.5$~KB on
the headline \texttt{grown\_capped\_dream} arm; full table in \S4.4). We
report deployment storage in the BF16 baseline because that is the honest
device-side accounting; an FP32 baseline overstates the dream-archive win.

\paragraph{Why ternary fails on a frozen-L0 substrate.}
Ternary quantization ($2$~bits/weight) on the frozen-Gaussian L0
projection costs $0.31$ absolute full-softmax accuracy on the smoke test:
random-Gaussian weights have no concentration of magnitude near zero, so
$2$-bit quantization throws away most of the routing structure that L1
and the head learned to depend on. Int8 ($\approx 1/127 \approx 0.8\%$
per-weight noise) preserves the routing intact. Unlocking ternary would
require quantization-aware training during the dream phase, or restricting
ternary to layers whose weights are concentrated at zero (a learned,
pruned L1 + head) --- both deferred to future work.

\paragraph{Epigenetic analogue.}
Phase 1 corresponds to terminal cell differentiation --- a cell that has
settled into its mature role no longer participates in plasticity. Phase 2
corresponds to compression of consolidated memory traces into
low-bandwidth long-term storage, freeing high-bandwidth substrate for new
learning.

\subsection{Mixed-Precision Substrate}
\label{sec:mixed-precision}

For deployment on resource-constrained devices the architecture supports a
mixed-precision mode in which weight Parameters $(W, b)$ are stored in
narrow precision (BF16) while all consolidation buffers ---
$\anchorW, \anchorB, F_W, F_b, \lambda, u, r, p$ --- remain in FP32. The
forward pass auto-casts inputs to the weight dtype, preserving the
callable interface; the EWC penalty and Fisher accumulation upcast cleanly
across the boundary.

This separation is necessary: pure-narrow-precision training degrades by
$\approx\!40\%$ due to optimizer-state precision loss in Adam's running
moments, but inference latency and memory cost are dominated by the
weights, which can safely run in BF16. The intended deployment pattern is
\emph{train in FP32 during sleep cycles, deploy in BF16 for always-on
inference}.

\subsection{The Ship-Wake-Extend Loop}
\label{sec:ship-wake-extend}

Deployment is not a one-shot training run that yields a fixed artifact; it
is an iterative loop in which the model is shipped, woken on new data,
allowed to grow plastic substrate, and re-shipped. We instantiate this
loop with a \emph{consolidation-dream pass} that fires once at a chosen
task boundary:
\begin{enumerate}[leftmargin=2em]
  \item \texttt{consolidation\_dream\_pass} --- full-coverage replay over
        all past tasks with archive-aware gradient masking (locks
        honored), plus one additional \texttt{archive\_block} call to
        catch rows that just settled.
  \item Optional permanent int8 snap of the archived rows (no restore
        step, since this is the shipping checkpoint).
  \item Cap lift:
        $\textit{current\_cap\_bytes} \leftarrow \textit{extension\_cap\_bytes}$.
        The network is allowed to grow new plastic substrate against the
        new task budget.
  \item The training loop continues over the new tasks with shared
        internal state --- manifold buffer, EWC anchors, archived rows,
        frustration counters all persist across the boundary.
\end{enumerate}

The result is a smaller plastic envelope in deployment than in training,
with new growth allocated only against the residual. A device-conscience
deployment cannot revisit the original training data: it has only what it
has stored (the manifold buffer, the archived weights) and what arrives in
the new data stream. The consolidation-dream pass turns shipping into a
controlled handoff: the network ``sleeps long'' before going to market,
locking in the rows it will never need to move again, then wakes on new
data in the field.

\subsection{Multi-Branch Absorption}
\label{sec:absorption}

The ship-wake-extend loop (\S\ref{sec:ship-wake-extend}) lets a single
organism continue learning across deployment cycles. A natural next
question is whether two organisms trained \emph{independently} on
disjoint task streams can be \emph{composed} into a single recipient
without retraining either of them. We construct a multi-branch
organism that does exactly this, with a shared frozen L0 random
projection as the only structural prerequisite.

\paragraph{Architecture.} A multi-branch organism comprises a shared
frozen L0 (the same Kaiming-initialized random projection used in
single-branch trioron, \S\ref{sec:methods}) and $N$ parallel
\emph{branches}. Each branch is a self-contained donor produced by an
independent training run: a frozen L1 substrate, a frozen head slice
over its own covered class subset, and the per-class manifold archive
$\{(\mu_c, \Sigma_c)\}_{c \in C_b}$ that is already computed for
manifold replay (\S\ref{sec:manifold-replay}). The branches share the same
L0 random projection by being born at the same seed; this is the
\emph{shared-seed invariant}, and is what makes donor weights
semantically valid in recipient code-space without distillation. We
enforce non-overlap of class coverage across branches:
$C_b \cap C_{b'} = \emptyset$ for $b \neq b'$.

\paragraph{Forward pass.} For input $x$:
\begin{enumerate}[leftmargin=2em]
  \item Project to code space: $z = \sigma(W_{L_0} x + b_{L_0})$ using
        the shared L0.
  \item For each branch $b$, compute the per-branch head logits
        $\ell^{(b)} = \text{branch}_b(z)$ over $C_b$.
  \item Compute per-branch generative log-likelihoods from the archive
        as a uniform mixture of per-class diagonal Gaussians:
        $$\log p_b(z) = \log\!\left(\tfrac{1}{|C_b|} \sum_{c \in C_b} \mathcal{N}(z;\mu_c, \Sigma_c)\right).$$
  \item Compute branch gates by softmax over branches:
        $$g_b(z) = \frac{\exp(\log p_b(z) / T)}{\sum_{b'} \exp(\log p_{b'}(z) / T)},$$
        where $T$ is a temperature ($T = 1$ default). Soft gates
        (rather than argmax) are deliberate: when archive
        log-likelihoods on a query are close, multiple branches
        co-fire and their log-probability contributions are blended,
        which is the regime in which absorption preserves union
        accuracy past the headline class.
  \item Compose into combined logits over the union
        $C_{\cup} = \bigsqcup_b C_b$ via the principled mixture-of-experts
        log-probability:
        $$\text{combined}_c(z) = \log g_{b(c)}(z) + \log\text{softmax}_c(\ell^{(b(c))}),$$
        where $b(c)$ is the unique branch covering $c$. Slots not covered
        by any branch are not in $C_{\cup}$.
\end{enumerate}

\begin{figure}[t]
\centering
\begin{tikzpicture}[
  >={Latex[length=2mm]},
  font=\small,
  L/.style={draw, rounded corners=3pt, align=center, inner sep=3pt, fill=white},
  frozen/.style={L, fill=gray!15},
  donor/.style={L, fill=blue!10, minimum width=30mm, align=center},
  combine/.style={L, fill=green!15, minimum width=98mm, align=center, inner sep=5pt},
  bias/.style={L, dashed, fill=violet!12, font=\scriptsize, align=center},
  s/.style={font=\scriptsize, align=center},
  arr/.style={->, thick, shorten >=1pt, shorten <=1pt}
]
\node[L] (in) {Input $x$};
\node[frozen, below=3mm of in, minimum width=68mm] (l0)
  {\textbf{Shared frozen L0}: $z = \sigma(W_{L_0} x + b_{L_0})$\\
   \scriptsize matched random seed across donors\\[-1pt]
   \scriptsize $\Rightarrow$ shared code space (shared-seed invariant)};
\draw[arr] (in) -- (l0);

\node[donor, below=14mm of l0, xshift=-38mm] (d1)
  {\textbf{Donor 1}\\\scriptsize $L_1^{(1)}$, head$^{(1)}$, classes $C_1$\\
   \scriptsize Archive $\{(\mu_c, \Sigma_c)\}_{c\in C_1}$};
\node[donor, below=14mm of l0]              (d2)
  {\textbf{Donor 2}\\\scriptsize $L_1^{(2)}$, head$^{(2)}$, classes $C_2$\\
   \scriptsize Archive $\{(\mu_c, \Sigma_c)\}_{c\in C_2}$};
\node[donor, below=14mm of l0, xshift=38mm] (d3)
  {\textbf{Donor 3}\\\scriptsize $L_1^{(3)}$, head$^{(3)}$, classes $C_3$\\
   \scriptsize Archive $\{(\mu_c, \Sigma_c)\}_{c\in C_3}$};

\draw[arr] (l0.south) -- (d1.north);
\draw[arr] (l0.south) -- (d2.north);
\draw[arr] (l0.south) -- (d3.north);

\node[combine, below=14mm of d2] (cmb) {%
  \textbf{Archive gate $+$ log-prob combine}\\[2pt]
  $g_b(z) = \mathrm{softmax}_b\bigl(\log p_b(z)/T\bigr)$,\;\;
  $\text{combined}_c(z) = \log g_{b(c)}(z) + \log\mathrm{softmax}_c\bigl(\ell^{(b(c))}\bigr)$\\[2pt]
  \scriptsize over union namespace $C_\cup = \bigsqcup_b C_b$ \;|\; no router training};

\draw[arr] (d1.south) -- node[s, sloped, above]{$\ell^{(1)}, \log p_1$} (cmb.north -| d1);
\draw[arr] (d2.south) -- node[s, right=1mm]{$\ell^{(2)}, \log p_2$} (cmb.north -| d2);
\draw[arr] (d3.south) -- node[s, sloped, above]{$\ell^{(3)}, \log p_3$} (cmb.north -| d3);

\node[bias, below=4mm of cmb, text width=86mm]
  (bias) {Optional per-class bias offset (dream-cycle calibration; trained on synthetic codes from union archive --- no real data)};
\draw[->, dashed, thick] (bias.north) -- (cmb.south);
\end{tikzpicture}
\caption{Multi-branch absorption (chimera) at $N=3$ donors. All
donors share the same frozen Gaussian L0 by being born at matching
random seeds, so donor weights are byte-valid in the recipient code
space (\emph{shared-seed invariant}). Each donor contributes per-class
logits $\ell^{(b)}$ and a manifold-archive log-likelihood
$\log p_b(z)$; the same archive that serves manifold replay
(\S\ref{sec:manifold-replay}) also serves as the analytic Expert-Gate
router. The combine operation places per-branch log-softmax
contributions into the union class namespace, weighted by the gate ---
no router training and no fusion-layer learning. The dashed per-class
bias offset is the optional dream-cycle calibration: it closes most of
the BTM-MoE full-softmax gap at a small task-aware cost (\S4.6).}
\label{fig:chimera}
\end{figure}

\paragraph{Why per-branch log-softmax.} Each donor's head was trained
under its own active-class mask, so absolute logit \emph{magnitudes}
across donors are not comparable. Combining raw logits weighted by gates
produces a biased argmax at full-softmax. Passing each branch's logits
through a log-softmax restricted to its own coverage normalizes the
contributions to a common log-probability scale, after which the
mixture combine is well-posed. This is an inference-only fix; no
parameters are added.

\paragraph{Why the mechanism is zero-shot.} Donor weights are byte-valid
in the recipient because L0 is shared (matched random projection
$\Rightarrow$ matched code distribution), and donor logits are placed
into recipient class-namespace via per-class indexing. The gate is
computed analytically from the archive --- a generative
\emph{Expert-Gate}-style router \citep{aljundi2017expert} using the
manifold buffer trioron already produces. No router training, no
fusion-layer learning, no calibration pass; pure paste-and-go.

\paragraph{Optional dream-cycle calibration.} For deployments where the
full-softmax metric is load-bearing, a tiny additive bias offset can be
trained to compensate for residual cross-donor magnitude mismatch. We
sample synthetic codes $z \sim \mathcal{N}(\mu_c, \Sigma_c)$ from the
union archive, forward through the assembled organism, and apply
cross-entropy on the global label. Two variants:
\begin{itemize}[leftmargin=1.5em]
  \item \emph{per-branch} ($N$ scalars): one bias broadcast uniformly
        over each donor's covered slots. Mathematically task-aware-safe
        (a uniform per-donor shift is monotonic on within-task argmax
        because active classes within a chained-15 task always belong
        to a single donor by disjoint coverage).
  \item \emph{per-class} ($|C_{\cup}|$ scalars): one bias per union
        class. Higher capacity, but \emph{can} perturb within-task
        argmax; trades a small task-aware loss for a larger full-softmax
        gain.
\end{itemize}
The dream cycle uses only the donors' own archives --- no real-data
retraining is performed.

\subsection{Experiment Design}
\label{sec:experiment-design}

\paragraph{Curriculum (chained-15).}
A 30-class class-incremental benchmark drawn from three datasets in
sequence:
\begin{itemize}[leftmargin=1.5em]
  \item 10 MNIST digit tasks (one new digit class per task);
  \item 10 Fashion-MNIST tasks (one new garment class per task);
  \item 10 EMNIST-letters tasks A..J (one new letter class per task).
\end{itemize}
Each task introduces exactly one new class. At evaluation time the network
is asked to classify across all classes seen so far (full-softmax) or
restricted to the same task or domain (task-aware, domain-aware). The
class-incremental framing forces the network to commit to absolute class
identities --- there is no task ID at inference --- and is the standard
hard regime for continual-learning evaluation.

\paragraph{Why class-incremental, not task-incremental.}
Task-incremental benchmarks supply the task ID at inference, which
collapses the problem to learning a per-task classifier.
Class-incremental does not, and is therefore the regime where
partition-style methods (PackNet, HAT) suffer their characteristic
full-softmax collapse. Reporting class-incremental from the start avoids
confusing the comparison.

\paragraph{Extension curriculum (chained-23).}
chained-15 above, plus 8 additional EMNIST-letters tasks K..R introduced
at the extension boundary; total 23 class-incremental tasks across 38
classes. Used only in the extension experiment of \S4.5.

\paragraph{Network configuration.}
\begin{itemize}[leftmargin=1.5em]
  \item Three layers: input $\to$ L0 $\to$ L1 $\to$ head.
  \item L0: $784 \to 128$, frozen-Gaussian random projection (weights
        drawn from $\mathcal{N}(0, 2/784)$ at init and never updated). See
        \S\ref{sec:manifold-replay} for why L0 is frozen.
  \item L1: $128 \to 32$ initial, grown adaptively up to
        $\textit{cap\_bytes} = 8\,\text{K}$ trainable parameters during
        chained-15, lifted to 16~K at the extension boundary.
  \item Head: $L_1\text{\_width} \to n_{\text{classes\_so\_far}}$, grown
        adaptively as new classes appear.
  \item Activations: ReLU on L1, identity on the head.
  \item Optimizer: Adam, $\text{lr} = 3 \times 10^{-3}$.
  \item Epochs: 4 per task. Loss: cross-entropy on the active label set.
  \item EWC strength: $\beta = 1000$ during waking and dream-replay
        (matched).
\end{itemize}

\paragraph{Trioron arms.}
\begin{itemize}[leftmargin=1.5em]
  \item \emph{fixed\_ewc\_small} --- no growth; L1 fixed at the
        chained-15 cap. Isolates the EWC + manifold-replay contribution at
        matched parameters.
  \item \emph{grown\_capped\_no\_dream} --- growth under EWC, no
        dream-time consolidation. Isolates the dream contribution.
  \item \emph{grown\_capped\_dream} --- growth + dream within cap. The
        default deployment arm.
  \item \emph{grown\_uncapped\_dream} --- growth past cap with manifold
        replay. The upper-bound headline arm.
\end{itemize}

\paragraph{Competitor families (5).}
We benchmark against one method from each major continual-learning
family:
\begin{itemize}[leftmargin=1.5em]
  \item \emph{Hippocampal buffer (exemplar rehearsal).} $K$ samples per
        past class stored. $K \in \{1, 10, 20, 50\}$.
  \item \emph{PackNet} \citep{mallya2018packnet}. Iterative magnitude
        pruning with per-task masks. \emph{PackNet matched} matches our
        trainable-parameter budget; \emph{PackNet standard} uses the
        original recipe.
  \item \emph{HAT} \citep{serra2018hat}. Hard binary attention masks per
        task, gating each layer. \emph{Matched} and \emph{standard}
        variants as above.
  \item \emph{Online EWC} \citep{schwarz2018progress}. Fisher EMA across tasks
        with $\gamma = 0.95$; learning-rate and $\beta$ tuned via grid
        search on the no-dream loss.
  \item \emph{LwF + EWC} \citep{li2018lwf, kirkpatrick2017overcoming}.
        Logit-distillation toward the anchored network on old-class
        columns, in addition to the EWC penalty.
\end{itemize}

\paragraph{Metrics.}
\begin{itemize}[leftmargin=1.5em]
  \item \emph{Full-softmax accuracy} (full): mean across all classes seen
        so far of top-1 accuracy with the unrestricted softmax. The hard
        class-incremental metric.
  \item \emph{Domain-aware accuracy} (domain): top-1 accuracy restricted
        to classes that share a dataset domain (digits, fashion, letters)
        with the query. Measures preservation of coarse-grained
        discrimination.
  \item \emph{Task-aware accuracy} (task): top-1 accuracy restricted to
        the classes of the \emph{task} the query belongs to. The
        pairwise-discrimination metric; the right one for trioron's
        deployment role of context-gated personalization atop a frozen
        LLM.
  \item \emph{Forgetting} (forget): mean across past tasks of (final-task
        accuracy on that task) $-$ (accuracy at the time the task was
        first introduced); positive values indicate forgetting, negative
        values indicate \emph{negative forgetting} (accuracy improves over
        time after the task ended).
  \item \emph{$\sigma$-vs-baseline}: paired-by-seed standardized effect
        size across $N = 3$ random seeds, reported alongside absolute
        $\Delta$.
\end{itemize}

\paragraph{Storage accounting.}
We report deployment storage in two columns: BF16 baseline (the honest
device-side number, since BF16 is the inference dtype) and FP32 baseline
(the training-time number). The manifold buffer is counted at FP32
(per-class $\mu$ and $\sigma$, $2 \cdot L_0 \cdot 4 = 1024$ bytes/class).
All headline storage numbers in \S4 use the BF16 baseline.

\paragraph{Reproducibility.}
All benches run on commodity CPU. Each seed is fully deterministic given
the seed value, modulo a known torch nondeterminism in archive-trigger
timing (see \S4.4 for the v1/v2 reconciliation). Launcher scripts under
\texttt{experiments/} accept explicit seed lists; bench logs and CSVs from
every reported run are committed to the public repository.

\bibliographystyle{plainnat}
\bibliography{refs}

\end{document}
